Conversas Hiperligadas: Novo Paradigma de Comunicação e Colaboração, potenciado pela Tecnologia WebRTC

### 2015-04-06 tese/template_project_latex.tex
[ADD] almost finished XMPP

\usepackage{acronym}
\newacro{API}{Application Programming Interface}
\newacro{BOSH}{Bidirectional-streams Over Synchronous HTTP}
\newacro{CSS}{Cascading Style Sheets}
\newacro{HTML}{HyperText Markup Language}
\newacro{HTTP}{Hypertext Transfer Protocol}
\newacro{ICE}{Interactive Connectivity Establishment}
\newacro{IPv4}{Internet Protocol Version 4}
\newacro{IPv6}{Internet Protocol Version 6}
\newacro{IP}{Internet Protocol}
\newacro{MPEG}{Moving Picture Experts Group}
\newacro{NAT}{Network Address Translation}
\newacro{PSTN}{Public Switched Telephone Network}
\newacro{SAMI}{Synchronized Accessible Media Interchange}
\newacro{SDP}{Session Description Protocol}
\newacro{SIP}{Session Initiation Protocol}
\newacro{SMIL}{Synchronized Multimedia Integration Language}
\newacro{SMS}{Short Message Service}
\newacro{SRT}{SubRip Text}
\newacro{STUN}{Session Traversal Utilities for NAT}
\newacro{SVG}{Scalable Vector Graphics}
\newacro{TCP}{Transmission Control Protocol}
\newacro{TMN}{This Means Nothing}
\newacro{TURN}{Traversal Using Relays around NAT}
\newacro{VoIP}{Voice Over IP}
\newacro{WebRTC}{Web Real-Time Communication}
\newacro{XML}{eXtensible Markup Language}
\newacro{XMPP}{Extensible Messaging and Presence Protocol}
\overfullrule=2cm

@@

  A technique called long polling consists on making the server hold the request until there is fresh information or expiring after some time, after the receival, the client makes another request. This results on a better network usage and a faster server response, but both simple polling and long polling requests are sent with \ac{HTTP} headers, which are data overkill, specialy for small messages.

  WebSockets provide bidirectional communications trough a full-duplex socket channel. WebSocket handshake phase specifies an \ac{HTTP} header in order to upgrade to websocket type of communication, the remainder messages are done without \ac{HTTP} headers, which leads to much smaller messages and better network usage.

@@

  %https://bloggeek.me/siganling-protocol-webrtc/
  {\color{red}[SIP]}

  \ac{SIP} can be used as a signaling method, it allows web browsers to have audio, video and \ac{SMS} capabilities like mobile phones. For instance, it's possible to interoperate web communications with \ac{SIP}, mobile and fixed phones.

  \ac{SIP} is used in \ac{VoIP} applications due to its compatibility with \ac{PSTN}. Service providers are disponibilizing access to their \ac{SIP} infrastrucures through WebSockets. Frameworks like jsSIP, QoffeeSIP and sipML5 are used on client side to parse and encode \ac{SIP} messages, making \ac{SIP} accessible to web based applications universe.

  {\color{red}[XMPP]}

  \ac{XMPP}'s bi-directional communication is achieved through \ac{BOSH}, which basically consists on long polling.

>>>>

\usepackage{acronym}
\newacro{API}{Application Programming Interface}
\newacro{BOSH}{Bidirectional-streams Over Synchronous HTTP}
\newacro{CSS}{Cascading Style Sheets}
\newacro{HTML}{HyperText Markup Language}
\newacro{HTTP}{Hypertext Transfer Protocol}
\newacro{ICE}{Interactive Connectivity Establishment}
\newacro{IPv4}{Internet Protocol Version 4}
\newacro{IPv6}{Internet Protocol Version 6}
\newacro{IP}{Internet Protocol}
\newacro{MPEG}{Moving Picture Experts Group}
\newacro{NAT}{Network Address Translation}
\newacro{PSTN}{Public Switched Telephone Network}
\newacro{RFC}{Request For Comments}
\newacro{SAMI}{Synchronized Accessible Media Interchange}
\newacro{SDP}{Session Description Protocol}
\newacro{SIP}{Session Initiation Protocol}
\newacro{SIMPLE}{SIP for Instant Messaging and Presence Leveraging Extensions}
\newacro{SMIL}{Synchronized Multimedia Integration Language}
\newacro{SMS}{Short Message Service}
\newacro{SRT}{SubRip Text}
\newacro{STUN}{Session Traversal Utilities for NAT}
\newacro{SVG}{Scalable Vector Graphics}
\newacro{TCP}{Transmission Control Protocol}
\newacro{TMN}{This Means Nothing}
\newacro{TURN}{Traversal Using Relays around NAT}
\newacro{VoIP}{Voice Over IP}
\newacro{WebRTC}{Web Real-Time Communication}
\newacro{XEP}{XMPP Extensions}
\newacro{XML}{eXtensible Markup Language}
\newacro{XMPP}{Extensible Messaging and Presence Protocol}
\overfullrule=2cm

@@

  A technique called long polling consists on making the server hold the request until there is fresh information or expiring after some time, after the receival, the client makes another request. This results on a better network usage and a faster server response, but both simple polling and long polling requests are sent with \ac{HTTP} headers, which adds data overhead, specialy for small messages.

  WebSockets provide bidirectional communications trough a full-duplex socket channel. WebSocket handshake phase specifies an \ac{HTTP} header in order to upgrade to websocket type of communication, the remainder messages are done without \ac{HTTP} headers, which leads to much smaller messages and better network usage. WebSockets may not be avalailable on every web browser, frameworks like \textit{socket.io}\footnote{\url{http://socket.io/}} and \textit{SockJS}\footnote{\url{http://github.com/sockjs}} uses \ac{HTTP} when there is no support for WebSockets.

@@

  \ac{SIP} is protocol used for negotiation, creation, modification and finalization of communication sessions between users. \ac{SIP} follows a client/server architecture with \ac{HTTP} like messages and it can be used as signaling protocol. The advantage of \ac{SIP} is the ability to make video and voice calls applications over \ac{IP} networks.

  The work group \ac{SIMPLE} proposed the creation of \ac{SIP} extensions, namely presence information (rfc5263\footnote{\url{http://tools.ietf.org/html/rfc5263}}) and instant messaging (rfc3428\footnote{\url{http://tools.ietf.org/html/rfc3428}}).

  \ac{SIP} is used in \ac{VoIP} applications due to its compatibility with \ac{PSTN}. Service providers are disponibilizing access to their \ac{SIP} infrastrucures through WebSockets. Frameworks like \textit{jsSIP}\footnote{\url{http://jssip.net/}}, \textit{QoffeeSIP}\footnote{\url{http://qoffeesip.quobis.com/}} and \textit{sipML5}\footnote{\url{http://sipml5.org/}} are used on client side to parse and encode \ac{SIP} messages, making \ac{SIP} accessible to web based applications universe.

  \ac{SIP} with WebSockets can be used as a signaling method for WebRTC applications, it allows web browsers to have audio, video and \ac{SMS} capabilities like mobile phones. For instance, it's possible to interoperate web communications with \ac{SIP} networks, mobile and fixed phones.

  \ac{XMPP} was initially developed for instant messaging (Jabber\footnote{\url{http://jabber.org/}}). It is nowadays an open technology for real-time communications.

  \ac{XMPP} messages are \ac{XML} based, which are attractive for applications that need structured messages and rich hypermedia. \ac{XMPP} advantage is the addition of extensions, for example XEP-0096\footnote{\url{http://xmpp.org/extensions/xep-0096.html}}, which adds file transfer capabilities between two entities and XEP-0045\footnote{\url{http://xmpp.org/extensions/xep-0045.html}} which enables multi-user chat.

  \ac{XMPP}'s bi-directional communication is achieved through \ac{BOSH} (XEP-0206\footnote{\url{http://xmpp.org/extensions/xep-0206.html}}), which basically consists on long polling. This kind of communication is also possible through WebSockets (rfc7395\footnote{\url{http://tools.ietf.org/html/rfc7395}}).

  Amongst multiple XMPP servers softwares are: ejabberd\footnote{\url{http://jabberd.im/}}, Metronome\footnote{\url{http://lightwitch.org/metronome}}, Openfire\footnote{\url{http://igniterealtime.org/projects/openfire/}} and Prosody\footnote{\url{http://prosody.im/}}. Ejabberd is the one that implements more \ac{RFC} specifications and \ac{XEP}s\footnote{\url{http://en.wikipedia.org/wiki/Comparison_of_XMPP_server_software}}.
